### 2018-06-27 abstract.tex
intagrating some gonuke comments

\begin{abstract}

This work compares and evaluates different computational approaches for modeling
off-normal operation of a gas centrifuge enrichment cascade.

The goal of this work focuses on developing the necessary understanding of
potential misuse of enrichment cascades. The latter would allow more effective
international safeguards designs and approaches.  While it is straightforward to
design an enrichment cascade under ideal conditions as a function of the
theoretical feed, product and tails assays, it is very difficult to find
reliable information about the behavior of a given cascade when the feed assay
does not match the design value. Several methods have been developed to assess
the behavior of an enrichment cascade in such circumstances. In addition of the
cut ($\theta$) those methods evaluate the feed to product, feed to tails and the
product to tails enrichment ratio, respectively, $\alpha$, $\beta$ and $\gamma$,
as a function of the cascade feed assay. As those four parameters depend on each
other, determining two of them allows to fully define the other.  The first
approach consists of fixing $\theta$ and $\alpha$ recomputing the corresponding
assays at each stages of the cascade. The second one maintains the ideal
condition of the cascade ($\alpha$ and $\beta$ fixed across the whole cascade),
modifying $\theta$ value at each stage accordingly. Both approaches have been
implemented into the Cyclus fuel cycle simulator\cite{cyclus, mbmore.2018}. The
third fixes $\theta$ and $\gamma$, using both $\alpha$ and $\beta$ at each stage
as free parameters. The third method has been investigated in
\cite{walker.2017}.

Following a description of each method and an evaluation of differences between
each approach, this work compares the results produced by these methods within
scenarios involving misuse of enrichment cascades simulated using the dynamic
nuclear fuel cycle simulator, Cyclus.

\end{abstract}