### 2018-06-27 experiement.tex

\section{The experiment}
This works focuses on comparing the different misuse models to a reference
calculation in which a single large cascade is build and designed to directly
produce \gls{HEU} from natural uranium.
This works uses the Cyclus fuel cycle simulator to allow material exchange
between facilities. The enrichment cascade algorithm have been implemented in
the $mbmore$ package \cite{mbmore.2018}.
In each cases, 5060 centrifuges have been used and spread across up to 30
different gas enrichment cascades.

\subsection{The cascade configuration}

\subsubsection{reference}

As mentioned, all the further calculations will be compared to the most
favorable configuration to produce \gls{HEU}, where all the available
centrifuges are used in a single large cascade designed to directly produce
\gls{HEU} from natural uranium, with a tails assay close to $0.3w\%$. The design
characteristic of the reference cascade are summarized in Table
\ref{tab:cascade_config}.

\subsubsection{default cascade}

The default cascade is the cascade design for normal civilian enrichment
operation, enriching natural uranium to about $3.5w\%$, with a tails assay close
to $0.3w\%$. This cascade will be layered and fed with uranium at higher
enrichment to evaluate the possibility to used them, with few or no tuning, to
produce \gls{HEU}. The characteristics of the default cascade are summarized in
Table \ref{tab:cascade_config}.

\begin{table}[htb]
\centering
  \caption{Summary of cascade design.}
\begin{tabular}{cccc}
\toprule

Cascade Design &       & Reference   & Default    \\
\midrule
Targeted  & Feed       & $0.71w\%$   & $0.71w\%$  \\
Assays    & Product    & $90w\%$     & $3.5w\%$   \\
          & Tails      & $0.3w\%$   & $0.3w\%$   \\
\midrule
Effective & Product    & $90.35w\%$  & $4.13w\%$  \\
Assays    & Tails      & $0.29w\%$   & $0.29w\%$  \\ 
\midrule
Stages    & Enriching  & 4          & 4          \\
Number    & Stripping  & 39         & 10         \\
\bottomrule
\end{tabular}
  \label{tab:cascade_config}
\end{table}

\subsection{scenarios}
In the following, cascades can be connected in tandem, where each set of cascade
in parallel is called a ``level``, as illustrated on Figure
\ref{fig:cascade_level}.
Seven different simulations have been simulated, to evaluate the effectiveness
of an enrichment cascade when used outside of its designed scope :
\begin{itemize}
\item one as the reference calculation, with a single centrifuges designed to
    directly produce \gls{HEU} from natural uranium,
\item three calculations (one per misuse model) were default cascade are
    chained to produce \gls{HEU}, without recycling the tails of each cascade,
    the cascade tails are directly sent to the waste,
\item three calculations (one per misuse model) were default cascade are
    chained to produce \gls{HEU}, and the tails of each cascade are recycled, blending
    the tails of one level in the feed of the previous level of
    cascades (see Figure \ref{fig:cascade_level}).
\end{itemize}

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.45]{flow}
    \caption{Schematic representation of the chained cascades with three levels,
    with the feed, product and the tails flows, respectively in blue, green and
    red. The dashed orange line represent the alternative tails flow when
    tails recycling is considered.}
    \label{fig:cascade_level}
\end{figure}

\subsection{Level population}
In order to assign the optimum number of cascades to each level, a virtual cut as
been computed as:
\begin{equation}
    \theta^{v}_{j} = \frac{F_{j}-N''_{j}}{N'_{j}-N''_{j}},
\end{equation}
where, $j$ represents a level of cascade and $N_{j}$, $N'_{j}$ and $N''_{j}$
respectively the feed, product and tails assay of the cascades at this level.

A flow equation similar to \eqref{eq:flow} is then solve to get the optimum
number of cascade per level. When the tails are not recycled, the $(1-\theta)$
terms are removed from the flow equation.  The results of the level population
are summarized in Table \ref{tab:level}.

As it is not possible to assign a fraction of an enrichment cascade, cascade per
level are rounded up for each level but the first one. To the first level is
attributed the remaining available cascades.

\subsection{Timestep effect}

As it can be observed on Figures \ref{fig:case1}, \ref{fig:case2} and
\ref{fig:case3}, the different cascade levels required up to $2n+1$ timesteps
before it starts to enrich uranium, where $n$ corresponds to the level number.
This corresponds to the time required for the material to propagate from the
natural uranium source to the n-th level, regardless of the duration of the
timestep (hour, day, month,...). This is also true for the number of timesteps
required to reach the equilibrium assays value in the tails reprocessing cases.
As timestep can be artificially small or large, one will only be considering
equilibrium values flow rates and assays in the following analysis.