### 2018-06-28 dis_clusion.tex
finish with gonuke comments

\section{Discussion}

We can observe that when the cascade is left completely untouched (Model A) or
when it is slightly retuned to maintain the tails to product enrichment factor
as well as the cut of each centrifuges (Model C), chaining the cascade we can
observe high increase of the enrichment at each level.  On the contrary, when
retuning the cut of each centrifuges to maintain the ideal state of the cascades
(Model B) while chaining them, the \gls{HEU} production rate is favored over the
enrichment gain.

The tails recycling allows each model to achieve a large gain in productivity,
even for then model B in which the number of levels required to reach $90w\%$ of
$^{235}$U in the uranium does not change. Even if no cascade chaining options
achieves the same production rate as direct enrichment, the model B with tail
recycling reached about $80\%$ of the optimum production rate. Such production
rate would allow the build up of a Significant Quantity of \gls{HEU} in less
than 8 months\ldots

\section{Conclusion and future works}

This works has investigated and quantify the difference between potential
retuning of a gaseous enrichment cascade in order to chain them to produce
\gls{HEU} initially tuned to produce uranium enrichment for commercial reactors.
One of this tuning method allows up to $80\%$ of the production rate of a single
large enrichment cascade designed specifically for \gls{HEU} production using
the same number of centrifuges.

This works will be extended to the near future with additional misuse method,
allowing for example the reconfiguration of the centrifuges in the cascades.

For this study, the usage of the Cyclus fuel cycle simulator was not really
required, it only allows a quick determination of the blending equilibrium. It
is planned to make use of the Cyclus Dynamical Resource Exchange full capability
in order to automatically assign the different cascades to the different level
as function of the resources availability, optimising the productions rates in
each cases.

While mathematically correct, the authors do not guaranty the feasibility
different misuse tuning methods implemented and are welcoming any insight on the
matter.