### 2019-11-20 motivation.tex

\section{Motivation}

Gas centrifuge cascades are usually designed to operate in an ideal manner, with
no losses in separative work. To achieve such ideal configuration, the cascade
is designed to be fed with a specific feed assay and produce the target
enrichment while rejecting tails at a fixed assay.

With the current international tensions regarding enrichment capabilities, this
work aims to measure the effectiveness of a symmetric enrichment cascade when
used outside of its designed scope and quantify the attractiveness of such way
to build up significant quantity of \gls{HEU}.

The present work investigates the performance ofs an enrichment cycle when chaining
gas enrichment cascades tuned for low enrich uranium production from natural
uranium to instead produce HEU. As literature on the mater is for obvious reason
limited, three behavior models have been implemented and used to evaluate the
response of an enrichment cascade when fed with different assays than designed one.
This work also takes advantage of the Cyclus\cite{cyclus} fuel cycle capabilities
to evaluate the assay values at equilibrium.